---
title: "Module 4 Challenge 4 List Production"
author: "Zach C"
date: "September 23, 2017"
output: html_document
---
#**Create a list representing a simplified primate taxonomy**
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
> Platyrrhini <-c("Cebidae","Atelidae","Pitheciidae")
> Platyrrhini
[1] "Cebidae"     "Atelidae"    "Pitheciidae"

> Catarrhini <-c("Cercopithecidae","Hylobatidae","Hominidae")
> Catarrhini
[1] "Cercopithecidae" "Hylobatidae"     "Hominidae"      
> 
> Anthropoidea<-list("Platyrrhini","Catarrhini")
> Anthropodiea
Error: object 'Anthropodiea' not found
> Anthropoidea
[[1]]
[1] "Platyrrhini"

[[2]]
[1] "Catarrhini"

> Platyrrhini <-list("Cebidae","Atelidae","Pitheciidae")
> Platyrrhini
[[1]]
[1] "Cebidae"

[[2]]
[1] "Atelidae"

[[3]]
[1] "Pitheciidae"

> Catarrhini <-list("Cercopithecidae","Hylobatidae","Hominidae")
> Catarrhini
[[1]]
[1] "Cercopithecidae"

[[2]]
[1] "Hylobatidae"

[[3]]
[1] "Hominidae"

> Anthropoidea<-list("Platyrrhini","Catarrhini")
> Anthropoidea
[[1]]
[1] "Platyrrhini"

[[2]]
[1] "Catarrhini"

> Platyrrhini
[[1]]
[1] "Cebidae"

[[2]]
[1] "Atelidae"

[[3]]
[1] "Pitheciidae"

> Anthropoidea
[[1]]
[1] "Platyrrhini"

[[2]]
[1] "Catarrhini"
```
>#Note, when first constructing the sublist, you can't put the names in quotes if they weren't originally defined with quotations. You'll have to go back and rename them later

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
> Anthropoidea<-list(Platyrrhini,Catarrhini)
> Anthropoidea
[[1]]
[[1]][[1]]
[1] "Cebidae"

[[1]][[2]]
[1] "Atelidae"

[[1]][[3]]
[1] "Pitheciidae"


[[2]]
[[2]][[1]]
[1] "Cercopithecidae"

[[2]][[2]]
[1] "Hylobatidae"

[[2]][[3]]
[1] "Hominidae"


> names(Anthropoidea)<-c("Platyrrhini,Catarrhini")
> Anthropoidea
$`Platyrrhini,Catarrhini`
$`Platyrrhini,Catarrhini`[[1]]
[1] "Cebidae"

$`Platyrrhini,Catarrhini`[[2]]
[1] "Atelidae"

$`Platyrrhini,Catarrhini`[[3]]
[1] "Pitheciidae"


$<NA>
$<NA>[[1]]
[1] "Cercopithecidae"

$<NA>[[2]]
[1] "Hylobatidae"

$<NA>[[3]]
[1] "Hominidae"


> names(Anthropoidea)<-c("Platyrrhini","Catarrhini")
> Anthropoidea
$Platyrrhini
$Platyrrhini[[1]]
[1] "Cebidae"

$Platyrrhini[[2]]
[1] "Atelidae"

$Platyrrhini[[3]]
[1] "Pitheciidae"


$Catarrhini
$Catarrhini[[1]]
[1] "Cercopithecidae"

$Catarrhini[[2]]
[1] "Hylobatidae"

$Catarrhini[[3]]
[1] "Hominidae"
```

> #Note: Can construct properly labeled Lists from sub lists by defining the content of those sublists, placing them within the greater list function and then renaming the sublists using the function names("greater List")<-c("sublist name 1", "sublist name 2"...)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
> 
> Tarsiodea <-list(Tarsiidae)
Error: object 'Tarsiidae' not found
> Tarsiodea <-list("Tarsiidae")
> Haplorhini<-;list(Tarsiodea,Anthropoidea)
Error: unexpected ';' in "Haplorhini<-;"
> Haplorhini<-list(Tarsiodea,Anthropoidea)
> Haplorhini
[[1]]
[[1]][[1]]
[1] "Tarsiidae"


[[2]]
[[2]]$Platyrrhini
[[2]]$Platyrrhini[[1]]
[1] "Cebidae"

[[2]]$Platyrrhini[[2]]
[1] "Atelidae"

[[2]]$Platyrrhini[[3]]
[1] "Pitheciidae"


[[2]]$Catarrhini
[[2]]$Catarrhini[[1]]
[1] "Cercopithecidae"

[[2]]$Catarrhini[[2]]
[1] "Hylobatidae"

[[2]]$Catarrhini[[3]]
[1] "Hominidae"



> names(Haplorhini)<-c("Anthropoidea","Tarsioidea")
> Haplorhini
$Anthropoidea
$Anthropoidea[[1]]
[1] "Tarsiidae"


$Tarsioidea
$Tarsioidea$Platyrrhini
$Tarsioidea$Platyrrhini[[1]]
[1] "Cebidae"

$Tarsioidea$Platyrrhini[[2]]
[1] "Atelidae"

$Tarsioidea$Platyrrhini[[3]]
[1] "Pitheciidae"


$Tarsioidea$Catarrhini
$Tarsioidea$Catarrhini[[1]]
[1] "Cercopithecidae"

$Tarsioidea$Catarrhini[[2]]
[1] "Hylobatidae"

$Tarsioidea$Catarrhini[[3]]
[1] "Hominidae"
```
>#Note, list titles were reversed, so this needed to be fixed

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
> names(Haplorhini)<-c("Tarsioidea","Anthropoidea")
> 
> 
> Haplorhini
$Tarsioidea
$Tarsioidea[[1]]
[1] "Tarsiidae"


$Anthropoidea
$Anthropoidea$Platyrrhini
$Anthropoidea$Platyrrhini[[1]]
[1] "Cebidae"

$Anthropoidea$Platyrrhini[[2]]
[1] "Atelidae"

$Anthropoidea$Platyrrhini[[3]]
[1] "Pitheciidae"


$Anthropoidea$Catarrhini
$Anthropoidea$Catarrhini[[1]]
[1] "Cercopithecidae"

$Anthropoidea$Catarrhini[[2]]
[1] "Hylobatidae"

$Anthropoidea$Catarrhini[[3]]
[1] "Hominidae"



> Lorisoidea<-c("Lorisidae","Galagidae")
> Lorisoidea
[1] "Lorisidae" "Galagidae"
> Lemuroidea<-c("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
> Lemuroidea
[1] "Cheirogaleidae" "Lepilemuridae"  "Indriidae"      "Lemuridae"     
[5] "Daubentoniidae"
> Lorisoidea<-list("Lorisidae","Galagidae")
> Lemuroidea<-list("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
> Lemuroidea
[[1]]
[1] "Cheirogaleidae"

[[2]]
[1] "Lepilemuridae"

[[3]]
[1] "Indriidae"

[[4]]
[1] "Lemuridae"

[[5]]
[1] "Daubentoniidae"

> Lorisoidea<-list("Lorisidae","Galagidae")
> Lorisoidea
[[1]]
[1] "Lorisidae"

[[2]]
[1] "Galagidae"

> Strepsirhini <-list(Lorisoidea,Lemuroidea)
> Strepsirhini
[[1]]
[[1]][[1]]
[1] "Lorisidae"

[[1]][[2]]
[1] "Galagidae"


[[2]]
[[2]][[1]]
[1] "Cheirogaleidae"

[[2]][[2]]
[1] "Lepilemuridae"

[[2]][[3]]
[1] "Indriidae"

[[2]][[4]]
[1] "Lemuridae"

[[2]][[5]]
[1] "Daubentoniidae"


> names(Strepsirhini)<-c(Lorisoidea,Lemuroidea)
Error in names(Strepsirhini) <- c(Lorisoidea, Lemuroidea) : 
  'names' attribute [7] must be the same length as the vector [2]
> names(Strepsirhini)<-c("Lorisoidea","Lemuroidea")
> Strepsirhini
$Lorisoidea
$Lorisoidea[[1]]
[1] "Lorisidae"

$Lorisoidea[[2]]
[1] "Galagidae"


$Lemuroidea
$Lemuroidea[[1]]
[1] "Cheirogaleidae"

$Lemuroidea[[2]]
[1] "Lepilemuridae"

$Lemuroidea[[3]]
[1] "Indriidae"

$Lemuroidea[[4]]
[1] "Lemuridae"

$Lemuroidea[[5]]
[1] "Daubentoniidae"


> Haplorhini
$Tarsioidea
$Tarsioidea[[1]]
[1] "Tarsiidae"


$Anthropoidea
$Anthropoidea$Platyrrhini
$Anthropoidea$Platyrrhini[[1]]
[1] "Cebidae"

$Anthropoidea$Platyrrhini[[2]]
[1] "Atelidae"

$Anthropoidea$Platyrrhini[[3]]
[1] "Pitheciidae"


$Anthropoidea$Catarrhini
$Anthropoidea$Catarrhini[[1]]
[1] "Cercopithecidae"

$Anthropoidea$Catarrhini[[2]]
[1] "Hylobatidae"

$Anthropoidea$Catarrhini[[3]]
[1] "Hominidae"


> Primates <-list(Haplorhini,Strepsirhini)
> Primates
[[1]]
[[1]]$Tarsioidea
[[1]]$Tarsioidea[[1]]
[1] "Tarsiidae"


[[1]]$Anthropoidea
[[1]]$Anthropoidea$Platyrrhini
[[1]]$Anthropoidea$Platyrrhini[[1]]
[1] "Cebidae"

[[1]]$Anthropoidea$Platyrrhini[[2]]
[1] "Atelidae"

[[1]]$Anthropoidea$Platyrrhini[[3]]
[1] "Pitheciidae"


[[1]]$Anthropoidea$Catarrhini
[[1]]$Anthropoidea$Catarrhini[[1]]
[1] "Cercopithecidae"

[[1]]$Anthropoidea$Catarrhini[[2]]
[1] "Hylobatidae"

[[1]]$Anthropoidea$Catarrhini[[3]]
[1] "Hominidae"




[[2]]
[[2]]$Lorisoidea
[[2]]$Lorisoidea[[1]]
[1] "Lorisidae"

[[2]]$Lorisoidea[[2]]
[1] "Galagidae"


[[2]]$Lemuroidea
[[2]]$Lemuroidea[[1]]
[1] "Cheirogaleidae"

[[2]]$Lemuroidea[[2]]
[1] "Lepilemuridae"

[[2]]$Lemuroidea[[3]]
[1] "Indriidae"

[[2]]$Lemuroidea[[4]]
[1] "Lemuridae"

[[2]]$Lemuroidea[[5]]
[1] "Daubentoniidae"
```
**#Final List of Primate Phylogeny**
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
> names(Primates)<-c("Haplorhini","Strepsirhini")

> Primates
$Haplorhini
$Haplorhini$Tarsioidea
$Haplorhini$Tarsioidea[[1]]
[1] "Tarsiidae"


$Haplorhini$Anthropoidea
$Haplorhini$Anthropoidea$Platyrrhini
$Haplorhini$Anthropoidea$Platyrrhini[[1]]
[1] "Cebidae"

$Haplorhini$Anthropoidea$Platyrrhini[[2]]
[1] "Atelidae"

$Haplorhini$Anthropoidea$Platyrrhini[[3]]
[1] "Pitheciidae"

$Haplorhini$Anthropoidea$Catarrhini
$Haplorhini$Anthropoidea$Catarrhini[[1]]
[1] "Cercopithecidae"

$Haplorhini$Anthropoidea$Catarrhini[[2]]
[1] "Hylobatidae"

$Haplorhini$Anthropoidea$Catarrhini[[3]]
[1] "Hominidae"



$Strepsirhini
$Strepsirhini$Lorisoidea
$Strepsirhini$Lorisoidea[[1]]
[1] "Lorisidae"

$Strepsirhini$Lorisoidea[[2]]
[1] "Galagidae"


$Strepsirhini$Lemuroidea
$Strepsirhini$Lemuroidea[[1]]
[1] "Cheirogaleidae"

$Strepsirhini$Lemuroidea[[2]]
[1] "Lepilemuridae"

$Strepsirhini$Lemuroidea[[3]]
[1] "Indriidae"

$Strepsirhini$Lemuroidea[[4]]
[1] "Lemuridae"

$Strepsirhini$Lemuroidea[[5]]
[1] "Daubentoniidae"


